\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[margin=2cm]{geometry}
\usepackage{amsthm}
\usepackage{cite}
\usepackage{graphicx}
\title{\textbf{Vector Spaces}\\(A003)}
\author{Paul Valentine}
\date{18 March 2020}
\begin{document}
\maketitle
\begin{abstract}
  Vector spaces appear in many forms of mathematics. Indeed $\mathbb{R}$ forms a vector space. Vectors are elements of a vector space. Understanding what defines a vector space can lead to some suprising catagorisations of vectos that are not seen in undergraduate mathematics. This paper provides a first touch introduction to what defines a vector space and some common charactoristics of a vector space that are encounted in everyday mathematics and physics.
\end{abstract}
\section{Introduction}
\subsection{Mathematical Spaces}
A space is a set with some added structure. In some sense the space contains the geometry and the structure its algerbraic nature. Two basic types of space may be considered - Linear Spaces and Topographical Spaces. Linear spaces are algerbraic spaces and form vector spaces. Linear operations in linear spaces provide a notion of straight lines and ellipses but cannot define perpendicular lines or circles from ellipses as there is no definition of an inner product\cite{A004}. The dimension of a vector space is the number miximum number of linearly independent vectos that span the space.\\ \\
Topological spaces are analytical and give rise to the notion of continuous functions and differentiation. A metric space is a topological space on which a distances between points are defined.\\ \\
Vectors in a Euclidean space have a length and therefore a norm. These are normed linear spaces that are also topological metric spaces. A inner product space is endowed with a bilinear form $(\langle \cdot, \cdot \rangle : V \times V \to F)$\cite{A004} called an inner product. All inner product spaces are normed spaces. \\ \\
A Riemannian manifold is a smooth manifold with tangent spaces are inner product spaces. If we lose the positive definit nature of the norm then we have a seminorm delivering a pseudo Riemannian space. This is the form of space defined in special relativity by the Lorentz transforms.

\subsection{Vector Spaces}
A vector space is a collection of objects called \textbf{vectors} that can be scaled and added together in a linear way. It is therefore a linear space. The scaling and addition rules for a vector space form a set of axioms. Two operations are defined on this space $\cdot$ and $+$. This paper describes the basics of vector spaces.
\section{Operations $\cdot$ and $+$}
Let $V$ be an ordered set.
\label{operations}
\begin{equation}
  \label{add}
  +:V \times V \to V
\end{equation}
Equation \ref{add} defines the addition rule where $\times$ is the cartesian production\cite{A002} and $V$ is a set. For example.
\begin{equation}
  +:\big((x_1, y_1),(x_2,y_2)\big) = \big((x_1+x_2,y_1+y_2)\big)
\end{equation}
 With a \textbf{field}\cite{wiki:mathfield} $F$ we have.
\begin{equation}
  \cdot : F \times V \to V
\end{equation}
Where $\cdot$ represents scalar multiplication. If $F$ is a complex field then we have a complex vector space. If $F$ is a real number field then we have a real vector space. A field is a set on which addition, subtraction, multiplication and division are defined. That it is, it is an albelian group with $0$ as the additive identity and $1$ as the multiplicative identity. Example:
\begin{equation}
\forall a\in \mathbb{R}  \cdot:\big(a,(x_1,y_1)\big) = \big((ax_1,ay_1)\big)
\end{equation}
Examples of fields are $\mathbb{R}, \mathbb{C}, \text{ and } \mathbb{Q}$.\\ \\
Coupling the above we say that a vector space V is over a field F.
\section{Axioms}
For the set $V$ to form a vector space the operations $+$ and $\cdot$ must conform to the following 8 axioms:
\begin{align}
  \text{Associativity of addition: } &\vec{u}+(\vec{v}+\vec{w}) = (\vec{u}+\vec{v})+\vec{w}\\
  \text{Commutativity of addition: }&\hat{u}+\hat{v} = \vec{v}+\vec{u}\\
  \text{Identity element of addition: }&\exists\:\vec{0}\in V \mid \hat{v} + \vec{0} = \hat{v} \:\forall \: \hat{v}\in V\\
  \text{Inverse element of addition: } &\forall \: \vec{v}\in V\: \exists \: -\vec{v} \mid \vec{v} + -\vec{v}=\vec{0}\\
  \text{Compatibility of scalar and field multiplication: }& a(b \vec{v}) = (ab) \vec{v}\\
  \text{Identity element of scale multiplication: }& 1 \vec{v} = \vec{v}\\
  \text{Distributivity of scalar multiplicatoin w.r.t vector addition: } &a(\vec{v}+\vec{u})= a\vec{v}+a\vec{u}\\
  \text{Ditributivity of scalar multiplication w.r.t field addition: }&(a+b)\vec{v} = a\vec{v}+b\vec{v}
\end{align}
\section{Examples}
\subsection{Coordinates spaces}
The simplest form a vector space is a field itself giving n tuples $(a_1, a_2, \dots, a_n)$.
\subsection{Function Space}
Let V be a vector space over a field F and X be any set. Then for $f:X \to V$, the set of f forms a vector space over F. Typically here $F,V \in \mathbb{R}$ as a common example.
\begin{align}
  &f,g:X \to V:\forall x \in X, \forall a \in F\\
  &(f+g)(x)=f(x)+g(x)\\
  &(a\cdot f)(x)=a \cdot f(x)
\end{align}
If X is also a vector space the set of linear maps $X \to V$ form a vector space over F called $Hom(X,V)$. A special case is the set of linear functions $V \to F$ which form a \textbf{dual} vector space.\\ \\
\subsection{Integration \cite{wiki:integration}}
The collectoin of integrable functions form a vector space from the above.
\begin{align}
  &f \mapsto \int_a^b f(x) dx\\
  &\int_a^b (\alpha f + \beta g)(x) dx = \int_a^b (\alpha f(x)+\beta g(x)) dx=\alpha \int_a^b f(x)dx + \beta \int_a^b g(x)dx
\end{align}
Since the function itself forms a vector space $f(x) dx$ takes the function to the field over which the integration occurs. Since this is $I:V \to F$ it places $f(x)dx$ $(I)$ in the dual vector space.
\section {Basis Vectors}
A set of vectors $B \subseteq V$ (Vector space) over a field $F$ is a linearly independent set that spans $V$ \cite{wiki:basis}. Spans means is able to describe all elements of $V$. Practically these means that all other elements of $V$ can be written as linear function of the elements in $B$.
\begin{align}
  &\text{Linear}\implies \forall b \in B, \forall a \in F, \sum_{i=1}^n a_i b_i =0 \implies a_i = 0\\
  &\text{Spans}\implies \forall v \in V \: \exists \: a \in F, b \in B \mid v = \sum_{i=1}^n a_i b_i
\end{align}
As an example, for a vector space of $\mathbb{R}^2$ over field $\mathbb{R}$ the basis vectors are $(1,0)$ and $(0,1)$.
\\ \\
$\square$
\bibliography{bibrefs}
\bibliographystyle{ieeetr}
\end{document}
